---
title: "R Problem of the Month"
date: March 27, 2015
---


# Introduction
To introduce random variables to his elementary statistics students, Mr. Garza gives the following problem:

> Suppose a person throws a dart and hits a circular board of radius 9 inches. Let $Y$ denote the distance in inches from the point of impact to the top center of the board. Is $Y$ a discrete or a continuous random variable? What values does the continuous random variable $Y$ take on?

The answers are that $Y$ is a continous random variable and that it can take on any value between 0 and 18 inclusive.

The problem does not ask students to find probabilities involving $Y$. To do so would require calculating integrals or otherwise be very clever.  But if he teaches his students R he may be able to ask them to approximate probabilites involving $Y$ by using [Monte Carlo methods](http://www.countbayesie.com/blog/2015/3/3/6-amazing-trick-with-monte-carlo-simulations).

Thus, the problem posed is this: assuming that all points on the dartboard are equally likely to be hit with the dart, find $P(0< Y<9)$ using R or other software.


```{r include=FALSE}
calculate_Y <- function(x,y) {
  centerX = 0
  centerY = 9
  Y = sqrt((x - centerX)^2 + (y -centerY)^2)
  return(Y)
}
    
throws = 100000

xs = runif(throws, min=-9, max=9) # generate random x-coordinates
ys = runif(throws, min=-9, max=9) # generate random y-coordinates

in.circle <- xs^2 + ys^2 <= 81  # check to see if they're in circle

# discard points outside the circle
xs <- xs[in.circle]
ys <- ys[in.circle]

# find those points that give 
condition <- calculate_Y(xs, ys) > 9 
mean(condition)  # this gives the answer
```

```{r echo=FALSE, fig.height=6, fig.width=6}
plot(xs, ys, xlim=c(-9,9), ylim=c(-9,9), 
  asp=1,
  pch = ".",
  main = "The Dartboard",
  xlab="", ylab="",
  col=ifelse(condition,"black","blue"))
```
